---
title: "Task 3: Bootstrapped confidence intervals"
author: "Grace Kumaishi"
date: "3/2/2021"
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(boot)
```

### 1) Histogram of bootstrapped proportions:

```{r, fig.align = "center"}
# Practice creating a data frame using tribble (unnecessary but fun...):
nonbinary <- tribble(~id, ~response, ~binary_response, 
              1, "yes", 1,
              2, "yes", 1,
              3, "no", 0,
              4, "yes", 1,
              5, "no", 0,
              6, "yes", 1,
              7, "yes", 1,
              8, "yes", 1,
              9, "no", 0,
              10, "no", 0,
              11, "yes", 1,
              12, "yes", 1,
              13, "no", 0,
              14, "yes", 1,
              15, "yes", 1,
              16, "no", 0,
              17, "yes", 1,
              18, "yes", 1,
              19, "yes", 1,
              20, "yes", 1,
              21, "yes", 1,
              22, "no", 0,
              23, "no", 0,
              24, "yes", 1,
              25, "no", 0,
              26, "no", 0,
              27, "yes", 1,
              28, "yes", 1,
              29, "no", 0,
              30, "yes", 1,
              31, "no", 0,
              32, "no", 0,
              33, "no", 0,
              34, "yes", 1,
              35, "yes", 1,
              36, "yes", 1)

# Get a vector of responses:
response_vec <- nonbinary$binary_response

# Create a function that will calculate the proportions of each bootstrapped sample:
proportion_fun <- function (x, i) {mean(x[i])} # Since the responses are binary, we can use the mean function here to calculate proportions

# Create 10k bootstrap samples by re-sampling from the response vector, using proportion function to calculate the proportion of each:
response_boot10k <- boot(response_vec,
                         statistic = proportion_fun,
                         R = 10000)

# Check the output:
# response_boot10k

# View original proportion:
#response_boot10k$t0 

# View 10k results:
#response_boot10k$t

# Convert vector to data frame:
response_boot10k_df <- data.frame(bs_prop = response_boot10k$t)

# Histogram of original:
# ggplot(data = nonbinary, aes(x = binary_response)) +
  # geom_histogram()

# Histogram of 10k bootstrap sample proportions:
ggplot(response_boot10k_df, aes(x = bs_prop)) +
  geom_histogram(fill = "mediumpurple4") +
  theme_minimal() +
  labs(x = "Proportion",
      y = "Count")
```

**Figure 1:** Histogram of proportion of nonbinary/genderqueer students who have experienced/are experiencing negative, exclusionary and/or hostile conduct from others. Proportions were calculated using 10,000 bootstrapped samples from the original dataset of 36 student responses. (Data: UCSB Campus Climate Project 2014).

<br>

### 2) Confidence interval:

```{r}
# Confidence interval:
ci <- boot.ci(response_boot10k, conf = 0.95, type = "perc")
ci
```

**CI takeaways:** We expect that 95% of calculated confidence intervals will contain the actual population parameter. In this case, 95 out of 100 times, the actual proportion of nonbinary students experiencing negative conduct will fall within our confidence interval (here shown as ~.44 to ~.78).

### Citation:
[Rankin & Associates Consulting. "UCSB Campus Climate Project Final Report." 2014.](https://campusclimate.ucop.edu/_common/files/pdf-climate/ucsb-full-report.pdf)




